# Power train mix by brand {#sec-power-train-mix-by-brand}

```{r}
#| label: setup
#| include: false

source("./scripts/setup-libraries.R")
source("./scripts/setup-data-vehicles-on-road-snapshot.R")

my_caption <- glue("Plot: Daniel Moul\nData: Traficom tieto.traficom.fi, trafi2.stat.fi",
                   "\nSnapshot of vehicles in traffic on Finnish mainland on 2023-09-30",
                   "; limited to model years {min_model_year}-{max_model_year}")

brand_cutoff_n <- 10

```

```{r}
data_for_table <- dta_by_brand_power_train |>
  select(brand, power_train, count) |>
  summarize(n = sum(count),
            .by = c(brand, power_train)
  ) |>
  complete(brand, power_train, fill = list(n = 0)) |>
  mutate(brand = fct_reorder(brand, -n, sum)) |>
  arrange(brand, desc(power_train)) |>
  pivot_wider(names_from = "power_train",
              values_from = "n") |>
  clean_names() |>
  rename(petrol_electric = petrol_electric_plug_in_hybrid,
         diesel_electric = diesel_electric_plug_in_hybrid) |>
  mutate(subtotal = petrol_electric + diesel_electric + battery_electric) |>
  mutate(total = petrol + diesel + subtotal + petrol_cng + natural_gas_cng + petrol_ethanol + other,
         pct = subtotal / total) |>
  relocate(diesel_electric, .after = petrol_electric) |>
  relocate(c(subtotal, pct), .after = battery_electric)

```


The following tables with 2023-09-30 snapshot data provide information multiple ways to simplify answering particular questions as noted in each section.

## Automotive brands with at least some battery electric power train market share in Finland 

### Sorted by number of battery electric vehicles

Which helps to answer the questions

* What are the battery electric vehicles on the road in Finland?
* If I were to spot a battery electric vehicle in Finland, which brands would it likely be?

```{r}
#| label: tbl-power-train-by-brand-electric-table-1
#| tbl-cap: "Automotive brands with at least some electric power train market share in Finland"
#| column: page

data_for_table  |>
  arrange(desc(battery_electric)) |>
  filter(battery_electric > 0) %>%
  mutate(rank = nrow(.) - rank(battery_electric) + 1,
         brand = as.character(brand)) |>
  relocate(rank) |>
  gt(rowname_col = "brand") |>
  tab_options(table.font.size = 10) |>
  tab_header(md(glue("**Automotive brands with some hybrid and electric<br>power train market share in Finland**",
                     "<br>Vehicles in traffic 2023-09-30"))
  ) |>
  tab_source_note(md("*Source: TRAFICOM tieto.traficom.fi, trafi2.stat.fi*")) |>
  tab_spanner(columns = c(petrol_electric, diesel_electric, battery_electric, pct, subtotal),
              label = "hybrid and electric") |>
  cols_align(columns = brand,
             align = "left") |>
  fmt_percent(columns = pct,
              decimals = 1) |>
  fmt_number(columns = total,
             decimals = 0) |>
  sub_missing() |>
  grand_summary_rows(
    columns = c(petrol:subtotal, petrol_cng:total),
    fns = list(id = "brand", label = "Total", fn = "sum"),
    fmt = ~ fmt_number(., use_seps = FALSE,
                       decimals = 0)
  ) |>
  data_color(
    columns = battery_electric,
    palette = "grey95"
  )

```

<br>

### Sorted by percent of brand's vehicles that are hybrid or electric

Which helps to answer the question

* Which brands are succeeding (or not succeeding) in moving to hybrid or battery electric power trains?

Note that in this ranking, having a large "install base" of older petrol or hybrid vehicles counts against a brand.

```{r}
#| label: tbl-power-train-by-brand-electric-table-2
#| tbl-cap: "Automotive brands with at least some electric power train market share in Finland"
#| column: page

data_for_table  |>
  arrange(desc(pct)) |>
  filter(subtotal > 0) %>%
  mutate(rank = nrow(.) - rank(pct) + 1,
         brand = as.character(brand)) |>
  relocate(rank) |>
  gt(rowname_col = "brand") |>
  tab_options(table.font.size = 10) |>
  tab_header(md(glue("**Automotive brands with some hybrid and electric<br>power train market share in Finland**",
                     "<br>Vehicles in traffic 2023-09-30"))
  ) |>
  tab_source_note(md("*Source: TRAFICOM tieto.traficom.fi, trafi2.stat.fi*")) |>
  tab_spanner(columns = c(petrol_electric, diesel_electric, battery_electric, pct, subtotal),
              label = "hybrid and electric") |>
  cols_align(columns = brand,
             align = "left") |>
  fmt_percent(columns = pct,
              decimals = 1) |>
  sub_missing() |>
  grand_summary_rows(
    columns = c(petrol:subtotal, petrol_cng:total),
    fns = list(id = "brand", label = "Total", fn = "sum"),
    fmt = ~ fmt_number(., use_seps = FALSE,
                       decimals = 0)
  ) |>
  data_color(
    columns = pct,
    palette = "grey95"
  )

```

<br>

### Sorted by number of vehicles all power trains

Which helps to answer the question

* Which brands have the largest and smallest market share in Finland independent of power train technology?

```{r}
#| label: tbl-power-train-by-brand-electric-table-3
#| tbl-cap: "Automotive brands with at least some electric power train market share in Finland"
#| column: page

data_for_table  |>
  arrange(desc(total)) |>
  filter(total > 0) %>%
  mutate(rank = nrow(.) - rank(total) + 1,
         brand = as.character(brand)) |>
  relocate(rank) |>
  gt(rowname_col = "brand") |>
  tab_options(table.font.size = 10) |>
  tab_header(md(glue("**Automotive brands with some hybrid and electric<br>power train market share in Finland**",
                     "<br>Vehicles in traffic 2023-09-30"))
  ) |>
  tab_source_note(md("*Source: TRAFICOM tieto.traficom.fi, trafi2.stat.fi*")) |>
  tab_spanner(columns = c(petrol_electric, diesel_electric, battery_electric, pct, subtotal),
              label = "hybrid and electric") |>
  cols_align(columns = brand,
             align = "left") |>
  fmt_percent(columns = pct,
              decimals = 1) |>
  sub_missing() |>
  grand_summary_rows(
    columns = c(petrol:subtotal, petrol_cng:total),
    fns = list(id = "brand", label = "Total", fn = "sum"),
    fmt = ~ fmt_number(., use_seps = FALSE,
                       decimals = 0)
  ) |>
  data_color(
    columns = total,
    palette = "grey95"
  )

```

<br>


## Automotive brands WITHOUT hybrid or electric vehicles

### Sorted by number of vehicles all power trains

Which helps to answer the question

* Which brands (from the perspective of the Finnish passenger vehicle market) seemingly are not even trying to move to hybrid or battery electric power trains?

```{r}
#| label: tbl-power-train-by-brand-electric-table-5
#| tbl-cap: "Automotive brands WITHOUT any electric power train market share in Finland"
#| column: page

data_for_table  |>
  arrange(desc(total)) %>%
  filter(subtotal == 0 & total > 0) %>%
  mutate(rank = nrow(.) - rank(total) + 1,
         brand = as.character(brand)) |>
  relocate(rank) |>
  gt(rowname_col = "brand") |>
  tab_options(table.font.size = 10) |>
  tab_header(md(glue("**Automotive brands without hybrid or electric<br>power train market share in Finland**",
                     "<br>Vehicles in traffic 2023-09-30"))
  ) |>
  tab_source_note(md("*Source: TRAFICOM tieto.traficom.fi, trafi2.stat.fi*")) |>
  tab_spanner(columns = c(petrol_electric, diesel_electric, battery_electric, pct, subtotal),
              label = "hybrid and electric") |>
  cols_align(columns = brand,
             align = "left") |>
  fmt_percent(columns = pct,
              decimals = 1) |>
  sub_missing() |>
  grand_summary_rows(
    columns = c(petrol:subtotal, petrol_cng:total),
    fns = list(id = "brand", label = "Total", fn = "sum"),
    fmt = ~ fmt_number(., use_seps = FALSE,
                       decimals = 0)
  ) |>
  data_color(
    columns = total,
    palette = "grey95" #c("dodgerblue"),
    #alpha = 0.1
  )

```

<br>

## Automotive brands sorted by alphabetical order

Which simplifies finding the information by brand.

```{r}
#| label: tbl-power-train-by-brand-electric-table-4
#| tbl-cap: "Automotive brands with at least some market share in Finland"
#| column: page

data_for_table  |>
  arrange(as.character(brand)) |>
  filter(total > 0) %>%
  gt(rowname_col = "brand") |>
  tab_options(table.font.size = 10) |>
  tab_header(md(glue("**Automotive brands with some market share in Finland**",
                     "<br>Vehicles in traffic 2023-09-30"))
  ) |>
  tab_source_note(md("*Source: TRAFICOM tieto.traficom.fi, trafi2.stat.fi*")) |>
  tab_spanner(columns = c(petrol_electric, diesel_electric, battery_electric, pct, subtotal),
              label = "hybrid and electric") |>
  cols_align(columns = brand,
             align = "left") |>
  fmt_percent(columns = pct,
              decimals = 1) |>
  sub_missing() |>
  grand_summary_rows(
    columns = c(petrol:subtotal, petrol_cng:total),
    fns = list(id = "brand", label = "Total", fn = "sum"),
    fmt = ~ fmt_number(., use_seps = FALSE,
                       decimals = 0)
  )

```

<br>

