# Failure reasons {#sec-failure-reasons}

```{r}
#| label: setup
#| include: false

source("./scripts/setup-libraries.R")
source("./scripts/setup-data-inspection-failures.R")

library(tidytext) # for reorder_within() and scale_y_reordered()
library(ggbump)

```

The data set includes the most common reasons that inspections failed for most (`model_year`, `brand`, `model`). Some (`model_year`, `brand`, `model`) include the second-most and third-most common reasons. To be included as a first, second or third most common reason, it must have been recorded for at least 10% of the vehicles in that (`model_year`, `brand`, `model`).

## Most common reasons

```{r}
dta_reason_1_overall <- dta_working_set |>
  count(failure_reason_1, sort = TRUE, name = "n_failures") |>
  mutate(rank = row_number()) |>
  left_join(reasons_df,
            join_by(failure_reason_1 == reason)) |>
  mutate(reason = coalesce(reason_short, failure_reason_1)) |>
  select(reason, n_failures, rank)

dta_reason_1_overall_yearly <- dta_working_set |>
  mutate(vehicle_age = 2022 - model_year) |>
  summarize(n_failures = n(),
            .by = c(vehicle_age, failure_reason_1)
            ) |>
  arrange(vehicle_age, desc(n_failures)) |>
  mutate(rank = row_number(),
         .by = vehicle_age
         ) |>
  left_join(reasons_df,
            join_by(failure_reason_1 == reason)) |>
  mutate(reason = coalesce(reason_short, failure_reason_1)) |>
  select(vehicle_age, reason, n_failures, rank)

dta_reason_1_overall_weighted <- dta_working_set |>
  mutate(n_failures = round(inspection_count * failure_rate)) |>
  count(failure_reason_1, sort = TRUE, wt = n_failures, name = "n_failures") |>
  mutate(rank = row_number()) |>
  left_join(reasons_df,
            join_by(failure_reason_1 == reason)) |>
  mutate(reason = coalesce(reason_short, failure_reason_1)) |>
  select(reason, n_failures, rank)
  
dta_reason_1_overall_yearly_weighted <- dta_working_set |>
  mutate(vehicle_age = 2022 - model_year,
         n_failures = round(inspection_count * failure_rate)) |>
  count(vehicle_age, failure_reason_1, wt = n_failures, name = "n_failures") |>
  arrange(vehicle_age, desc(n_failures)) |>
  mutate(rank = row_number(),
         .by = vehicle_age
         ) |>
  left_join(reasons_df,
            join_by(failure_reason_1 == reason)) |>
  mutate(reason = coalesce(reason_short, failure_reason_1)) |>
  select(vehicle_age, reason, n_failures, rank)

dta_reasons_all_weighted <- dta_working_set  |>
  pivot_longer(cols = starts_with("failure_reason"),
               names_to = "failure_number",
               values_to = "reason") |>
  filter(reason != "All model years") |> # found in failure_reason_2; not helpful
  left_join(reasons_df,
            join_by(reason == reason)) |>
  mutate(reason = coalesce(reason_short, reason)) |>
  mutate(failure_rate_weight = case_match(
    failure_number,
    "failure_reason_1" ~ 1.0,
    "failure_reason_2" ~ 0.35,
    "failure_reason_3" ~ 0.15),
    n_failures = round(failure_rate_weight * inspection_count * failure_rate)
  ) |>
  #count(reason, sort = TRUE, wt = n_failures) |>
  summarize(n_failures = sum(n_failures),
            .by = reason) |>
  arrange(desc(n_failures)) |>
  mutate(rank = row_number())

dta_reasons_all_yearly_weighted <- dta_working_set |>
  pivot_longer(cols = starts_with("failure_reason"),
               names_to = "failure_number",
               values_to = "reason") |>
  filter(reason != "All model years") |> # found in failure_reason_2; not helpful
   left_join(reasons_df,
            join_by(reason == reason)) |>
  mutate(reason = coalesce(reason_short, reason)) |>
  mutate(vehicle_age = 2022 - model_year,
         failure_rate_weight = case_match(
           failure_number,
           "failure_reason_1" ~ 1.0,
           "failure_reason_2" ~ 0.35,
           "failure_reason_3" ~ 0.15),
         n_failures = round(failure_rate_weight * inspection_count * failure_rate)
  ) |>
  count(vehicle_age, reason, sort = TRUE, wt = n_failures, name = "n_failures") |>
  mutate(rank = row_number(),
         .by = vehicle_age
  )

```

Below I rank reasons by the number of inspection failures reported having this reason in two ways: (a) considering only the most common reason for each (`model_year`, `brand`, `model`); and (b) also including second- and third-most common reasons when they are available. I estimated the rate of second and third-most common reasons (see [@sec-all-reasons]). The two rankings are quite similar.

```{r}
#| label: tbl-reasons-overall
#| tbl-cap: "Reasons for inspection failure: all ages, brands and models"
#| tbl-subcap: 
#|   - "Most common reasons"
#|   - "All reasons"
#| layout-ncol: 2
##| column: page

dta_reason_1_overall_weighted |>
  gt()

dta_reasons_all_weighted |>
  gt()

```

<br>

## Most common reasons by vehicle age

I use `vehicle_age` instead of `model_year` below, since it is a more natural way to think about `model_year` when considering 2022 inspection data.

As vehicles age and are driven farther, the reasons for inspection failure change.

```{r}
#| label: fig-reasons-overall-yearly
#| fig-cap: "Most common reasons for inspection failure by vehicle age"
#| fig-width: 20
#| fig-height: 10
#| column: screen-right
#| warning: false

dta_reason_1_overall_yearly_weighted |>
  ggplot(aes(vehicle_age, rank, color = reason)) + #failure_reason_1),
  geom_bump(show.legend = FALSE) +
  geom_point(show.legend = FALSE) +
  geom_label(aes(label = reason), #failure_reason_1),
             show.legend = FALSE,
             size = 5) +
  scale_x_continuous(breaks = 1:15,
                     position = "top") +
  scale_y_reverse(breaks = 1:15) +
  theme(panel.border = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(size = 18),
        axis.title.x = element_text(size = 18, hjust = 0),
        axis.title.y = element_text(size = 18, hjust = 1)) +
  labs(
    title = "Most common reasons for inspection failure by vehicle age",
    subtitle = "Ranked by number of failures. All years, brands and models. Rank 1 is most common",
    x = "Vehicle age",
    y = "Rank",
    caption = my_caption
  )
```

<br>

The higher-ranked reasons are much more frequent, and as vehicles age, the top reasons occur even more often: the top reason is more than 15x at age 15 compared to age 4 (compared to rank 2 reason at age for, since the first reason is "not provided"). See @fig-reasons-overall-yearly above.

```{r}
#| label: fig-reasons-overall-yearly-frequency-scaled
#| fig-cap: "Frequency of most common reasons for inspection failure by vehicle age"
#| fig-width: 20
#| fig-height: 10
#| column: screen-right
#| warning: false

data_for_plot <- dta_reason_1_overall_yearly_weighted |>
  mutate(n_k = round(n_failures / 1000),
         plot_label = if_else(n_k == 0,
                              glue("{n_failures}"),
                              glue("{n_k}K"))
  )

data_for_label_left <- data_for_plot |>
  filter(vehicle_age == min(vehicle_age))

data_for_label_right <- data_for_plot |>
  filter(vehicle_age == max(vehicle_age))

data_for_plot |>
  ggplot(aes(vehicle_age, rank, color = reason)) +
  geom_bump(show.legend = FALSE) +
  geom_point(show.legend = FALSE) +
  geom_label(aes(label = plot_label, size = n_failures),
             show.legend = FALSE) + #size = 7
  geom_text(data = data_for_label_left,
            aes(x = vehicle_age, y = rank, color = reason, label = reason),
            hjust = 1, nudge_x = -0.4, size = 7,
            show.legend = FALSE) +
  geom_text(data = data_for_label_right,
            aes(x = vehicle_age, y = rank, color = reason, label = reason),
            hjust = 0, nudge_x = 0.5, size = 7,
            show.legend = FALSE) +
  scale_x_continuous(breaks = 4:15,
                     position = "top") +
  scale_y_reverse(breaks = 1:15) +
  scale_size_continuous(range = c(4, 11)) +
  coord_cartesian(x = c(2, 17)) +
  theme(panel.border = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(size = 18),
        axis.title.x = element_text(size = 18, hjust = 0),
        axis.title.y = element_text(size = 18, hjust = 1)) +
  labs(
    title = "Frequency of most common reasons for inspection failure by vehicle age",
    subtitle = "Ranked by number of failures. All years, brands and models. Rank 1 is most common.",
    x = "Vehicle age",
    y = "Rank",
    caption = my_caption
  )
```


<br>

## All reasons by vehicle age {#sec-all-reasons}

In some cases the data includes a second and third most common reason. For the following plots, I estimated the frequency of failures for second and third most common reasons for any (`vehicle_age`, `brand`, `model`) as follows:

* most common: `r percent(1.0)` * `failure_rate`
* second most common: `r percent(0.35)` * `failure_rate`
* third most common: `r percent(0.15)` * `failure_rate`

I then summed them by reason and plotted the results below (@fig-reasons-all-yearly and @fig-reasons-all-yearly-frequency-scaled).

```{r}
#| label: fig-reasons-all-yearly
#| fig-cap: "Reasons for inspection failure by vehicle age (including all reasons)"
#| fig-width: 20
#| fig-height: 10
#| column: screen-right
#| warning: false

dta_reasons_all_yearly_weighted |>
  ggplot(aes(vehicle_age, rank, color = reason)) +
  geom_bump(show.legend = FALSE) +
  geom_point(show.legend = FALSE) +
  geom_label(aes(label = reason),
             show.legend = FALSE,
             size = 5) +
  scale_x_continuous(breaks = 1:15,
                     position = "top") +
  scale_y_reverse(breaks = 1:15) +
  theme(panel.border = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(size = 18),
        axis.title.x = element_text(size = 18, hjust = 0),
        axis.title.y = element_text(size = 18, hjust = 1)) +
  labs(
    title = "Reasons for inspection failure by vehicle age (including all reasons)",
    subtitle = "Ranked by number of failures. All years, brands and models. Rank 1 is most common.",
    x = "Vehicle age",
    y = "Rank",
    caption = my_caption
  )
```

<br>

The higher-ranked reasons are much more frequent, and as vehicles age the top reasons occur more often: the top reason is more than 15x at age 15 compared to age 4 (the first reason listed in age 4 is no answer ("not provided"). See @fig-reasons-overall-yearly above.

```{r}
#| label: fig-reasons-all-yearly-frequency-scaled
#| fig-cap: "Frequency of all reasons for inspection failure by vehicle age"
#| fig-width: 20
#| fig-height: 10
#| column: screen-right
#| warning: false

data_for_plot <- dta_reasons_all_yearly_weighted |>
  mutate(n_k = round(n_failures / 1000),
         plot_label = if_else(n_k == 0,
                              glue("{n_failures}"),
                              glue("{n_k}K"))
  )

data_for_label_left <- data_for_plot |>
  filter(vehicle_age == min(vehicle_age))

data_for_label_right <- data_for_plot |>
  filter(vehicle_age == max(vehicle_age))

data_for_plot |>
  ggplot(aes(vehicle_age, rank, color = reason)) +
  geom_bump(show.legend = FALSE) +
  geom_point(show.legend = FALSE) +
  geom_label(aes(label = plot_label, size = n_failures),
             show.legend = FALSE) + #size = 5
  geom_text(data = data_for_label_left,
            aes(x = vehicle_age, y = rank, color = reason, label = reason),
            hjust = 1, nudge_x = -0.4, size = 7,
            show.legend = FALSE) +
  geom_text(data = data_for_label_right,
            aes(x = vehicle_age, y = rank, color = reason, label = reason),
            hjust = 0, nudge_x = 0.5, size = 7,
            show.legend = FALSE) +
  scale_x_continuous(breaks = 4:15,
                     position = "top") +
  scale_y_reverse(breaks = 1:15) +
  scale_size_continuous(range = c(4, 11)) +
  #expand_limits(x = c(2, 17)) +
  coord_cartesian(x = c(2, 17)) +
  theme(panel.border = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_text(size = 18),
        axis.title.x = element_text(size = 18, hjust = 0),
        axis.title.y = element_text(size = 18, hjust = 1)) +
  labs(
    title = "Frequency of all reasons for inspection failure by vehicle age",
    subtitle = "Ranked by number of failures. All years, brands and models. Rank 1 is most common.",
    x = "Vehicle age",
    y = "Rank",
    caption = my_caption
  )
```

<br>

## Brand differences in most common reasons

***TODO: Pick a few brands and show their different results***

***TODO: Make a plot function***


```{r}
knitr::knit_exit()
```

